---
title: "Class Project on Using Knit on Activity Data"
output: html_document
---


## R Markdown



# Code for reading in the dataset and/or processing the data


```{r}
##library(ggplot2)
##library(dplyr)

MyData <- read.csv("activity.csv",header=TRUE,sep=",",fill=TRUE)

# check to see
head(MyData)
```

# Histogram of the total number of steps taken each day
For this part of the assignment, we are ignoring the missing values in the dataset. Cleaning the data first.
{r}
MyData$Date <- mapply(function(x) as.POSIXct(x,"%Y-%m-%d"),MyData$date) 
head(MyData)

Calculate the total number of steps taken per day
```{r}
SumDate <- aggregate(steps ~ date, rm.na=TRUE, data= MyData, FUN=sum)
head(SumDate, n=20)
hist(SumDate$steps,main="Total steps per day",xlab="Steps")
```

# Mean and median number of steps taken each day

```{r}
MM <- as.integer(mean(SumDate$steps))
print(paste("The mean of the total steps taken each day is ",MM))
MD <- as.integer(median(SumDate$steps))
print(paste("The median of the total steps taken each day is ",MD))
```

# Time series plot of the average number of steps taken

```{r}
x <- aggregate(steps~interval,data=MyData, na.rm=TRUE, FUN=mean)
head(x)
plot(x,type="l", xlab = "Intervals of 5-min", ylab="Total steps")

```

# The 5-minute interval that, on average, contains the maximum number of steps
```{r}
MS <- max(x$steps)
IN <- x$interval[x$steps == MS]

print(paste("The max steps ", MS, " is the 5 min interval # ", IN))
```

# Code to describe and show a strategy for imputing missing data

```{r}
str(MyData)
summary(MyData)
SS <- sum(is.na(MyData))
print(paste("The total missing values in dataset is ",SS))

MyNewData = transform(MyData, steps = ifelse(is.na(steps), mean(steps, na.rm=TRUE), steps))
```
# Histogram of the total number of steps taken each day after missing values are imputed

```{r}
NewSumDate <- aggregate(steps ~ date, rm.na=TRUE, data= MyNewData, FUN=sum)
MM1 <- as.integer(mean(NewSumDate$steps))
print(paste(" The mean steps for the original data is ",MM, " and the mean for the imputed data is ",MM1))
MD1 <- as.integer(median(NewSumDate$steps))
print(paste(" The mean steps for the original data is ",MD, " and the mean for the imputed data is ",MD1))

hist(NewSumDate$steps,main="Total steps per day with Imputed Data",xlab="Steps")
```
# Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}


wends <- c("Saturday","Sunday")
#logical vector
weekendYes <- weekdays(as.Date(MyData$date)) %in% wends
#initialize all to weekday first
MyNewData$DayType <- "weekday"
#now set the weekends
MyNewData$DayType[weekendYes] <- "weekend"
MyNewData$DayType <- factor(MyNewData$DayType)

WKDAY <- aggregate(steps ~ interval,  data= subset(MyNewData, DayType=="weekday"), mean)
WKEND <- aggregate(steps ~ interval,  data= subset(MyNewData, DayType=="weekend"), mean)
head(WKDAY)
head(WKEND)
#make 2 panel plot
par(mfrow=c(2,1),mar=c(5,4,4,1))
plot(WKDAY,type="l",xlab="Weekday 5 min interval",ylab="Average steps")
plot(WKEND,type="l",xlab="Weekend 5 min interval",ylab="Average steps")

```





